<!-- Set PArameters Module -->
<!-- Set the chnkPrefix to make all chunks unique in the overall folder -->
```{r, echo=TRUE, eval=TRUE, warning=FALSE}
chnkPrefix <- "B.9.pca."
VersionPdfExt <- VersionPdfExt <- paste0(".",chnkPrefix,"V", gsub("-", "", Sys.Date()), ".pdf")

library(gplots)
library(RColorBrewer)
library(lattice)
library(genefilter)
library(tidyr)
library(ggplot2)
```


```{r, echo=F, eval=TRUE, warning=FALSE, results=F}
dfPCA <- Obio@dfPCA
```


```{r pca_data_prep, echo=TRUE, eval=TRUE, warning=FALSE, results=F}

###############################################################################
## Add PCA plot                                                              ##
chnkVec <- as.vector(NULL, mode = "character")
plotList <- list()
tag <- "PCAvariationPerDimension"


if (exists("Obio")){
    pos <- grep("PCApercentVar", slotNames(Obio))
    if (!is.null(Obio@PCApercentVar)){
        PCApercentVar <- Obio@PCApercentVar
    }
} else {
    PCApercentVar <- NULL
}

plotList[[tag]] <- ggplot(
        data = dfPCA,
        aes(x=PC1, y=PC2, fill = sample_group)
    ) + geom_vline(xintercept = 0, color = "grey", size=0.5
    ) + geom_hline(yintercept = 0, color = "grey", size=0.5
    ) + geom_point(
        size=2,
        shape = 21
    ) 

if (!is.null(PCApercentVar)){
    plotList[[tag]] <- plotList[[tag]] + labs(
        title = "PCA Plot", 
        x = paste0("PC1 \n ",round(100* Obio@PCApercentVar[1]),"% variability explained"),
        y = paste0("PC2 \n ",round(100* Obio@PCApercentVar[2]),"% variability explained")
    )
} else {
    plotList[[tag]] <- plotList[[tag]] + labs(
        title = "PCA Plot", 
        x = paste0("PC1"),
        y = paste0("PC2")
    )
}

plotList[[tag]] <- plotList[[tag]] +  theme_bw() + theme(
        axis.text.y   = element_text(size=8),
        axis.text.x   = element_text(size=8),
        axis.title.y  = element_text(size=12),
        axis.title.x  = element_text(size=12),
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        plot.title = element_text(hjust = 0.5, size = 12)
        #) + scale_fill_manual(values=c("#999999", "#E69F00")
    )


###########################################################################
## Save plot to file                                                     ##
FNbase <- paste0("PCA12", VersionPdfExt)
FN <- paste0(Obio@parameterList$reportFigDir, FNbase)
FNrel <- paste0("report_figures/", FNbase)
FNrelT <- paste0("report_tables/", FNbase)
    
pdf(FN)
    print(plotList[[tag]])
dev.off()
##                                                                       ##
###########################################################################
    


link <- paste0('<a href="https://biologic.crick.ac.uk/',Obio@parameterList$project_id,'/pca?x_axis=PC1&y_axis=PC2', '" target="_blank">here</a>')

figCap <- paste0(
    "**Figure ",
    figureCount,
    ":** Variation in the first two PCA Dimensions. Download a pdf of this figure [here](", FNrel, "). ",
    "Further PCA dimensions are available interacively ", link, ". " 
)
 
figureCount <- figureCount + 1
    
NewChnk <- paste0(
paste0("### PCA_Plot \n"),
            "\n```{r ReferencePCA1, results='asis', echo=F, eval=TRUE, warning=FALSE, fig.cap='",figCap,"'}\n",
            "\n",
            "\n print(plotList[['",tag,"']])",
            "\n cat(  '\n')",
            "\n\n\n```\n"   
)

chnkVec <- c(
    chnkVec,
    NewChnk
)

            

## Done with PCA plot                                                        ##
###############################################################################


```


```{r variation_data_prep, echo=TRUE, eval=TRUE, warning=FALSE, results=F}

###############################################################################
## Add Variation estimate plot                                               ##




if (length(unique(dfDesign$dataseries)) > 1){
    independentDesignColSector <- c(
        "dataseries"
    )
} else {
    independentDesignColSector <- as.vector(NULL, mode = "character")
}

if (length(dfDesign$sample.group) > 1){
    independentDesignColSector <- c(
        independentDesignColSector,
        "sample.group"
    )
}

if (length(dfDesign$replicate) > 1){
    independentDesignColSector <- c(
        independentDesignColSector,
        "replicate"
    )
}

pos <- grep("^f_", names(dfDesign))
if (length(pos) > 0){
  independentDesignColSector <- c(
    names(dfDesign)[grep("^f_", names(dfDesign))]
  )

} else if (length(unique(dfDesign$dataseries)) > 1){
    independentDesignColSector <- c(
      "dataseries"
    )
}

###############################################################################
## Create independent variations plot                                        ##

designColSelector = unique(c(independentDesignColSector, "sample.id"))    

if (length(unique(dfDesign$sample.id)) > 42) {
    rld <- vst(Obio@ObjDds)    
} else {
    rld <- rlog(Obio@ObjDds)    
}
rv = rowVars(assay(rld))

## Select most variable genes
select = order(rv, decreasing = TRUE)[seq_len(length(mostVarIDs))]
dfTemp = t(assay(rld)[select, ])

pc <- prcomp(dfTemp, center=TRUE, scale=FALSE)

colDatMin = unique(dfDesign[, designColSelector])
rownames(colDatMin) = as.vector(colDatMin$sample.id)

colDatMin$sample.id <- NULL
#colnames(colData)[1] = "condition"

###############################################################################
## Get PCA Loadings                                                          ##

dfBase <- t(dfTemp)
pcaGenes = prcomp(scale(dfBase))

dfPcaGenes = data.frame(pcaGenes$x)
dfPcaGenes[[ alignmentGeneID]] <- row.names(dfPcaGenes)
#         
dfLoad <- pcaGenes$rotation
dfLoad <- t(dfLoad)
#         
dfBase <- data.frame(dfBase)
#         
#######################################################################
## helper function                                                   ##

determinePCAloadings <- function(
    lmFitDim = 'lmFitDim',
    dfBase = 'dfBase',
    primary.alignment.gene.id = 'primary.alignment.gene.id'
){
    for (i in 1:ncol(dfBase)){
        corVar <- names(dfBase)[i]
        regression_formula <- as.formula(paste0("pcaGenes$x[,",lmFitDim,"]~", corVar))
        fit <- lm(regression_formula, data=dfBase)
        #summary(fit)
        p.value <- summary(fit)$coefficients[corVar,"Pr(>|t|)"]
        estimate <- summary(fit)$coefficients[corVar,"Estimate"]
        intercept <- summary(fit)$coefficients[grep("Intercept", row.names(summary(fit)$coefficients)),"Estimate"]
        rsquared <- summary(fit)$r.squared

        new.row <- data.frame(p.value, estimate, intercept, rsquared)

        row.names(new.row) <- corVar
        if (i!=1){
            dfRes <- rbind(dfRes, new.row)
        } else {
            dfRes <- new.row
        }
    }

    dfRes <- dfRes[order(dfRes$p.value, decreasing = FALSE),]
    dfRes[["padj"]] <- p.adjust(dfRes$p.value, method = "BH")
    #names(dfRes) <- paste0(names(dfRes), ".PCA", lmFitDim)
    dfRes[[primary.alignment.gene.id]] <- row.names(dfRes)


    ## Add log10p column ##
    padj  <- names(dfRes)[grep("padj", names(dfRes))]
    lg10p <- gsub("padj", "lg10p", padj)

    for (z in 1:length(padj)){
        preprocess <- as.numeric(dfRes[,padj[z]])

        if (length(grep("padj", padj[i])) > 0){
            preprocess <- as.numeric(res[,padj[z]])
            minNum <- min(preprocess[preprocess != 0])
            preprocess[preprocess == 0] <- minNum
        } else {
            preprocess <- as.numeric(dfRes[,padj[z]])
        }

        temp <- -1*log10(preprocess)
        #temp[temp >= 50] = 50
        dfRes[,lg10p[z]] <- temp
    }
    dfRes[["PCAdimension"]] <- paste0("PCA", lmFitDim)
    ## Done adding log10p ##

    return(dfRes)
}

## helper function                                                   ##
#######################################################################

pcaDimensionsToInvestigate = c(1:6)

for (l in 1:length(pcaDimensionsToInvestigate)){
    dfTemp <- determinePCAloadings(
        lmFitDim = pcaDimensionsToInvestigate[l],
        dfBase = dfBase,
        primary.alignment.gene.id = alignmentGeneID
    )

    if (l ==1){
        dfAll <- dfTemp
    } else {
        dfAll <- rbind(
            dfAll,
            dfTemp
        )
    }
    print(paste0(l, " dimensions out of ", length(pcaDimensionsToInvestigate), " processed."))
}

names(dfAll) <- gsub("intercept.PCA", "intercept_PCA", names(dfAll))
names(dfAll) <- gsub("estimate.PCA", "contrast_P_PCA_estimatePCA", names(dfAll))
names(dfAll) <- gsub("padj.PCA", "contrast_P_padj_PCA", names(dfAll))
names(dfAll) <- gsub("lg10p.PCA", "contrast_P_lg10p_PCA", names(dfAll))
names(dfAll) <- gsub("rsquared.PCA", "r2_PCA", names(dfAll))

Obio@dfPCAgenes <- dfAll
#         
#         ## Adding plots ##
#         obj <- createPCAloadingPlots(obj)

## Doen get PCA loadings                                                     ##
###############################################################################


covar_PC_frame <- rbind(
    data.frame(
        Component=1:(nrow(pc$x)-1),
        spread(
            data.frame(
                v=names(colDatMin),
                val=NA_real_
            ),
            key=v, 
            value=val
        )
    )
)


tag <- "independentVariation"

#if (mode == "independentVariation"){        
    covar_PC_frame <- covar_PC_frame[c("Component", names(colDatMin))]
            for (i in 1:nrow(covar_PC_frame)) {
                ## old code from gavin below ##
                fit <- lm(pc$x[,i]~., data=colDatMin)
                covar_PC_frame[i,-1] <- drop1(fit, test="F")[names(covar_PC_frame)[-1],"Pr(>F)"]
                
                ## replaced 25032019 ##
                # Fit each variable individually @
            }
    
    
    plotFrame <- gather(covar_PC_frame, key=Covariate, value=p, -Component)
        plotFrame <- plotFrame[order(plotFrame$Component, decreasing = FALSE),]
        
        if (nrow(plotFrame) > 20) {
            plotFrame <- plotFrame[1:(length(names(colDatMin)) * 20), ]
        }
        
        ## Cut to 10 dimensions ##
        
        plotList[[tag]] <- ggplot(
            plotFrame, 
            aes(x=Component, y=Covariate, fill=-log10(p))) +
            geom_raster() +
            scale_fill_gradient(low="grey90", high="red") +
            theme_classic() + 
            coord_fixed() +
            scale_x_continuous( labels = unique(plotFrame$Component), breaks = unique(plotFrame$Component)
            )
        
        ###########################################################################
    ## Save plot to file                                                     ##
    FNbase <- paste0("Independent.variation.per.pca.dimension", VersionPdfExt)
    FN <- paste0(Obio@parameterList$reportFigDir, FNbase)
    FNrel <- paste0("report_figures/", FNbase)
    
    pdf(FN)
        print(plotList[[tag]])
    dev.off()
    ##                                                                       ##
    ###########################################################################
    
    figCap <- paste0(
        '**Figure ',
        figureCount,
        ':** Independent sources of Variation per principal component. ',
        'Download a pdf of this figure <a href="',FNrel,'" target="_blank">here</a>. '
    )
 
    figureCount <- figureCount + 1
   
    NewChnk <- paste0(
            "### Independent Source of Variation Per PCA Component ",
            "\n```{r var-per-pca-independent, results='asis', echo=F, eval=TRUE, warning=FALSE, fig.cap='",figCap,"'}\n",
            "\n",
            "\n print(plotList[['",tag,"']])",
            "\n cat(  '\n')",
            "\n\n\n```\n"   
    )

    chnkVec <- c(
        chnkVec,
        NewChnk
    )
        
###############################################################################
## Now the plot tolarating dependent variations                              ##

tag <- "dependentVariation"
    
dependentDesignColSelector<- c(
    independentDesignColSector,
    names(Obio@dfDesign)[grep("comp_", names(dfDesign))],
    names(Obio@dfDesign)[grep("LRT_", names(dfDesign))]
)    
    
covar_PC_frame <- rbind(
            data.frame(
                Component=1:(nrow(pc$x)-1),
                spread(
                    data.frame(
                        v=names(colDatMin),
                        val=NA_real_
                    ),
                    key=v, 
                    value=val
                )
            )
        )

mode <- "dependentVariation"
    ## Do fitting individually ##
    ## Check that all selVec entries exist
        fitVars <- names(covar_PC_frame)
        fitVars <- fitVars[fitVars != "Component"]
        covar_PC_frame <- covar_PC_frame[c("Component", names(colDatMin))]


        for (i in 1:nrow(covar_PC_frame)) {
            ## old code from gavin below ##
            for (j in 1:length(fitVars)){
                corVar <- fitVars[j]

                if (length(unique(dfDesign[, corVar])) > 1) {
                    pcDim <- paste0("pc$x[,",i,"]")
                    regressionFormula <- as.formula(paste(pcDim, corVar, sep="~"))
                    fit <- lm(regressionFormula, data=colDatMin)
                    pVal <- as.vector(summary(fit)$coefficients[,4][2])
                    covar_PC_frame[i, corVar] <- pVal
                }
            }
        }    
        
    plotFrame <- gather(covar_PC_frame, key=Covariate, value=p, -Component)
    plotFrame <- plotFrame[order(plotFrame$Component, decreasing = FALSE),]

    if (nrow(plotFrame) > 20) {
        plotFrame <- plotFrame[1:(length(names(colDatMin)) * 20), ]
    }
    
       
    ## Cut to 10 dimensions ##
    
    plotList[[tag]] <- ggplot(
        plotFrame, 
        aes(x=Component, y=Covariate, fill=-log10(p))) +
        geom_raster() +
        scale_fill_gradient(low="grey90", high="red") +
        theme_classic() + 
        coord_fixed() +
        scale_x_continuous( labels = unique(plotFrame$Component), breaks = unique(plotFrame$Component)
        ) +  ggplot2::theme(
        axis.text.y   = ggplot2::element_text(size=8),
        axis.text.x   = ggplot2::element_text(size=8),
        axis.title.y  = ggplot2::element_text(size=8),
        axis.title.x  = ggplot2::element_text(size=8),
        axis.line = ggplot2::element_line(colour = "black"),
        panel.border = ggplot2::element_rect(colour = "black", fill=NA, size=1),
        plot.title = ggplot2::element_text(hjust = 0.5, size = 12)
    ) + ggplot2::labs(title = "Independent Sources of Variation per PCA Component")
        
        ###########################################################################
    ## Save plot to file                                                     ##
    FNbase <- paste0("Dependent.permissive.variation.per.pca.dimension", VersionPdfExt)
    FN <- paste0(Obio@parameterList$reportFigDir, FNbase)
    FNrel <- paste0("report_figures/", FNbase)
    
    pdf(FN)
        plotList[[tag]]
    dev.off()
    ##                                                                       ##
    ###########################################################################
    
    figCap <- paste0(
        '**Figure ',
        figureCount,
        ':** Dependent-tolerant sources of Variation per principal component. ',
        'Download a pdf of this figure <a href="',FNrel,'" target="_blank">here</a>. '
    )
 
    figureCount <- figureCount + 1
   
    NewChnk <- paste0(
            "### Dependent-tolerant Source of Variation Per PCA Component ",
            "\n```{r var-per-pca-dependent-tolerant, results='asis', echo=F, eval=TRUE, warning=FALSE, fig.cap='",figCap,"'}\n",
            "\n",
            "\n print(plotList[['",tag,"']])",
            "\n cat(  '\n')",
            "\n\n\n```\n"   
    )

    chnkVec <- c(
        chnkVec,
        NewChnk
    )
       
    
    
## Done                                                                      ##
###############################################################################
        


## Add PCA loadings


    ## Add genes driving this PCA dimension ## 
#     if (!is.null(Obio@plotCollection$PCA1_PCA_fitting)){
#         
#         pFit <- Obio@plotCollection$PCA1_PCA_fitting
#         
#         
#         ###########################################################################
#         ## Save plot to file                                                     ##
#         FNbase <- paste0("Variation.per.pca.dimension.", VersionPdfExt)
#         FN <- paste0(Obio@parameterList$reportFigDir, FNbase)
#         FNrel <- paste0("report_figures/", FNbase)
#     
#         pdf(FN)
#             print(pFit)
#         dev.off()
#         ##                                                                       ##
#         ###########################################################################
#         link <- paste0(
#             "https://biologic.crick.ac.uk/",
#             Obio@parameterList$project_id,
#             "/scatterplot?x_axis=contrast_P_PCA_estimatePCA1&y_axis=contrast_P_lg10p_PCA1&highlight_gene=&cat_id=ag_lab_categories__10")
#     figCap <- paste0(
#         "**Figure ",
#         figureCount,
#         ":** Genes driving principal components. ",
#         "Download a pdf of this figure [here](", FNrel, "). ",
#         "Genes driving this - and other PCA dimensions can be accessed interactively [here](", link, "). " 
#     )
#  
#     figureCount <- figureCount + 1
#    
#     NewChnk <- paste0(
#             "### Genes Driving PCA Components ",
#             "\n```{r var-per-pca-genes, results='asis', echo=F, eval=TRUE, warning=FALSE, fig.cap='",figCap,"'}\n",
#             "\n",
#             "\n print(pFit)",
#             "\n cat(  '\n')",
#             "\n\n\n```\n"   
#     )
#     chnkVec <- c(
#         chnkVec,
#         NewChnk
#     )
#         
#     }
# }
## Done adding PCA plots                                                     ##
###############################################################################
if (length(plotList) > 2){
    tabVar <- ".tabset .tabset-fade .tabset-dropdown"
} else {
    tabVar <- ".tabset .tabset-fade .tabset-pills"
}

```

## Principal Component Analysis (PCA) {`r tabVar`}

A birds eye view of your data can be obtained by looking at the results of the principal component analysis (PCA). The principal component analysis looks at your count dataset as a whole and determines how 'close' two samples are in terms of overall data structure. First of all, you want your replicated to cluster together. After that, you will be able to determine how different various sets of sample groups are.

A more detailed explanation on PCA is give in this [PCA video](https://www.youtube.com/watch?v=_UVHneBUBW0&feature=youtu.be).

```{r pca_var_plot, echo=TRUE, eval=TRUE, warning=FALSE, results="asis"}
cat(paste(knitr::knit(text = chnkVec, quiet = T), collapse = '\n'))
```